---
title: "Ghosts"
author: "Paulo Sigrist"
date: "11/14/2016"
output: html_document
---

```{r setup, include=FALSE}

library('dplyr') # data manipulation
library('ggplot2') # visualization
library('ggthemes') # visualization
library('mice')
library('randomForest') # prediction
knitr::opts_chunk$set(echo = TRUE)
```

# Kaggle - Ghosts Competition

First, load data:

```{r}
# Load data
train <- read.csv('train.csv', stringsAsFactors = T)
test  <- read.csv('test.csv', stringsAsFactors = T)


theColors <- levels(factor(train$color))
# Train - 371 x 7
# Test  - 529 x 6

full <- bind_rows(train, test)

ghosts <- train[train$type == 'Ghost',]
ghouls <- train[train$type == 'Ghoul',]
goblins <- train[train$type == 'Goblin',]

```

Data fields

id - id of the creature
bone_length - average length of bone in the creature, normalized between 0 and 1
rotting_flesh - percentage of rotting flesh in the creature
hair_length - average hair length, normalized between 0 and 1
has_soul - percentage of soul in the creature
color - dominant color of the creature: 'white','black','clear','blue','green','blood'
type - target variable: 'Ghost', 'Goblin', and 'Ghoul'


## Simple Random Forrest

First attempt is to try to use Random Forrest and the type is based on the other features
```{r}
set.seed(754)


# 
rf_model <- randomForest(factor(type) ~ bone_length + rotting_flesh + hair_length + has_soul + color, data = train)

# Obter a importancia de cada var
importance <- importance(rf_model)
varImportance <- data.frame(Variables = row.names(importance), Importance = round(importance[,'MeanDecreaseGini'],2))

# Criar um rank
rankImportance <- varImportance %>% mutate(Rank = paste0('#', dense_rank(desc(Importance))))


ggplot(rankImportance, aes(x = reorder(Variables, Importance), 
    y = Importance, fill = Importance)) +
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
    hjust=0, vjust=0.55, size = 4, colour = 'red') +
  labs(x = 'Variables') +
  coord_flip() + 
  theme_few()
```

We can see that `hair_length` and `has_soul` are the two most important features. Let's see in a scatter plot:

```{r}
ggplot(train, aes(x = hair_length, y = has_soul, shape = type, color = type)) + geom_point()
```

As we can see, there is a concentration of ghosts in the left bottom part, a concentration of goblins in the middle, and ghouls in the top right of the chart.

The first conclusion is that ghosts have less hair and soul, ghouls have more hair and soul, and goblins are in the middle.

There are 4 creatures that are close to each other, two ghosts, a ghoul and a goblin, with close hair size e soul. What the difference between them? Let's find this 4 guys and see!

```{r}
creatures <- train[train$has_soul > 0.30 & train$has_soul < 0.375 & train$hair_length > 0.425 & train$hair_length < 0.475,]
creatures
ggplot(creatures, aes(x = hair_length, y = has_soul, shape = type, color = type)) + geom_point()

```


```{r}
# Bone length by color
ggplot(ghosts, aes(color, bone_length)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(ghouls, aes(color, bone_length)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(goblins, aes(color, bone_length)) + geom_boxplot() + geom_jitter(width = 0.2)

# Rotting flesh by color
ggplot(ghosts, aes(color, rotting_flesh)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(ghouls, aes(color, rotting_flesh)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(goblins, aes(color, rotting_flesh)) + geom_boxplot() + geom_jitter(width = 0.2)

# Hair length by color
ggplot(ghosts, aes(color, hair_length)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(ghouls, aes(color, hair_length)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(goblins, aes(color, hair_length)) + geom_boxplot() + geom_jitter(width = 0.2)

# Soul by color
ggplot(ghosts, aes(color, has_soul)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(ghouls, aes(color, has_soul)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(goblins, aes(color, has_soul)) + geom_boxplot() + geom_jitter(width = 0.2)


```

## First submission

Without further analysis, let's run the prediction with the Random Forrest and see the results!
```{r}

prediction <- predict(rf_model, test)

test$type <- prediction

ggplot(test, aes(x = hair_length, y = has_soul, shape = type, color = type)) + geom_point()
```

### Saving the results
```{r}
# Save the solution to a dataframe with two columns: PassengerId and Survived (prediction)
solution <- data.frame(id = test$id, type = test$type)

# Write the solution to file
write.csv(solution, file = 'rf_mod_Solution1.csv', row.names = F)
```

