---
title: "Ghosts"
author: "Paulo Sigrist"
date: "11/14/2016"
output: html_document
---

```{r setup, include=FALSE}

library('dplyr') # data manipulation
library('ggplot2') # visualization
library('ggthemes') # visualization
library('mice')
library('randomForest') # prediction
library('e1071')

knitr::opts_chunk$set(echo = TRUE)
```

# Kaggle - Ghosts Competition

First, load data:

```{r}
# Load data
train <- read.csv('train.csv', stringsAsFactors = T)
test  <- read.csv('test.csv', stringsAsFactors = T)


theColors <- levels(factor(train$color))
# Train - 371 x 7
# Test  - 529 x 6

full <- bind_rows(train, test)

ghosts <- train[train$type == 'Ghost',]
ghouls <- train[train$type == 'Ghoul',]
goblins <- train[train$type == 'Goblin',]

```

Data fields

id - id of the creature
bone_length - average length of bone in the creature, normalized between 0 and 1
rotting_flesh - percentage of rotting flesh in the creature
hair_length - average hair length, normalized between 0 and 1
has_soul - percentage of soul in the creature
color - dominant color of the creature: 'white','black','clear','blue','green','blood'
type - target variable: 'Ghost', 'Goblin', and 'Ghoul'


## Simple Random Forrest

First attempt is to try to use Random Forrest and the type is based on the other features
```{r}
set.seed(754)


# 
rf_model <- randomForest(factor(type) ~ bone_length + rotting_flesh + hair_length + has_soul + color, data = train)

# Obter a importancia de cada var
importance <- importance(rf_model)
varImportance <- data.frame(Variables = row.names(importance), Importance = round(importance[,'MeanDecreaseGini'],2))

# Criar um rank
rankImportance <- varImportance %>% mutate(Rank = paste0('#', dense_rank(desc(Importance))))


ggplot(rankImportance, aes(x = reorder(Variables, Importance), 
    y = Importance, fill = Importance)) +
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
    hjust=0, vjust=0.55, size = 4, colour = 'red') +
  labs(x = 'Variables') +
  coord_flip() + 
  theme_few()
```

We can see that `hair_length` and `has_soul` are the two most important features. Let's see in a scatter plot:

```{r}
ggplot(train, aes(x = hair_length, y = has_soul, shape = type, color = type)) + geom_point() + 
  geom_vline(xintercept = 0.42, color="red") + geom_vline(xintercept = 0.475, color="red") + 
  geom_hline(yintercept = 0.30, color="red") + geom_hline(yintercept = 0.375, color="red")
```

As we can see, there is a concentration of ghosts in the left bottom part, a concentration of goblins in the middle, and ghouls in the top right of the chart.

The first conclusion is that ghosts have less hair and soul, ghouls have more hair and soul, and goblins are in the middle.

There are 4 creatures that are close to each other, two ghosts, a ghoul and a goblin, with close hair size e soul. What the difference between them? Let's find this 4 guys and see!

```{r}
creatures <- train[train$has_soul > 0.30 & train$has_soul < 0.375 & train$hair_length > 0.425 & train$hair_length < 0.475,]
creatures
ggplot(creatures, aes(x = hair_length, y = has_soul, shape = type, color = type)) + geom_point()

```


```{r}
# Bone length by color
ggplot(ghosts, aes(color, bone_length)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(ghouls, aes(color, bone_length)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(goblins, aes(color, bone_length)) + geom_boxplot() + geom_jitter(width = 0.2)

# Rotting flesh by color
ggplot(ghosts, aes(color, rotting_flesh)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(ghouls, aes(color, rotting_flesh)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(goblins, aes(color, rotting_flesh)) + geom_boxplot() + geom_jitter(width = 0.2)

# Hair length by color
ggplot(ghosts, aes(color, hair_length)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(ghouls, aes(color, hair_length)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(goblins, aes(color, hair_length)) + geom_boxplot() + geom_jitter(width = 0.2)

# Soul by color
ggplot(ghosts, aes(color, has_soul)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(ghouls, aes(color, has_soul)) + geom_boxplot() + geom_jitter(width = 0.2)
ggplot(goblins, aes(color, has_soul)) + geom_boxplot() + geom_jitter(width = 0.2)


```


```{r}
ggplot(train, aes(type, bone_length)) + geom_boxplot() + labs(title="Type x Bone Length") + geom_jitter(width = 0.2)
ggplot(train, aes(type, rotting_flesh)) + geom_boxplot() + labs(title="Type x Rotting Flesh") + geom_jitter(width = 0.2)
ggplot(train, aes(type, hair_length)) + geom_boxplot() + labs(title="Type x Hair Length") + geom_jitter(width = 0.2)
ggplot(train, aes(type, has_soul)) + geom_boxplot() + labs(title="Type x Has Soul") + geom_jitter(width = 0.2)
```


## Outliers

Looking the data distribution, we can see a lot of outliers in the training data

### By color

As we saw, Hair Length and Soul are the features that mainly differantiate the creatures. Let's run a K-Means, with K = 3 (types of creatures) using this two features, and see the results

```{r}

train$colorAsInteger <- as.integer(train$color)
# Run K-Means for Hair and Soul fatures
theCluster <- kmeans(train[,c(4,5)], 3)
# Add the cluster into the test dataset
train$clusterByHairSoul <- theCluster$cluster

# See the results
ggplot(train, aes(x = hair_length, y = has_soul, shape = type, color = clusterByHairSoul)) + geom_point()


```

We can see the cluster is accurate, but we have some creatures classified as others, let's classify this creatures as outliers.

```{r}
table(train[train$type == 'Ghost',]$clusterByHairSoul)
table(train[train$type == 'Ghoul',]$clusterByHairSoul)
table(train[train$type == 'Goblin',]$clusterByHairSoul)

# None is an outlier
train$outlier <- FALSE

# Ghosts
clusterGhosths <- which(table(train[train$type == 'Ghost',]$clusterByHairSoul) == max(table(train[train$type == 'Ghost',]$clusterByHairSoul)))
train[train$type == 'Ghost' & train$clusterByHairSoul != clusterGhosths,]$outlier = TRUE

# Ghouls
clusterGhouls <- which(table(train[train$type == 'Ghoul',]$clusterByHairSoul) == max(table(train[train$type == 'Ghoul',]$clusterByHairSoul)))
train[train$type == 'Ghoul' & train$clusterByHairSoul != clusterGhouls,]$outlier = TRUE

# Goblins
clusterGoblins <- which(table(train[train$type == 'Goblin',]$clusterByHairSoul) == max(table(train[train$type == 'Goblin',]$clusterByHairSoul)))
train[train$type == 'Goblin' & train$clusterByHairSoul != clusterGoblins,]$outlier = TRUE

ggplot(train, aes(x = hair_length, y = has_soul, shape = type, color = outlier)) + geom_point()

```


## Second attempt

```{r}
rf_model <- randomForest(factor(type) ~ bone_length + rotting_flesh + hair_length + has_soul + color, data = train[train$outlier == FALSE,], proximity = TRUE)

# Obter a importancia de cada var
importance <- importance(rf_model)
varImportance <- data.frame(Variables = row.names(importance), Importance = round(importance[,'MeanDecreaseGini'],2))

# Criar um rank
rankImportance <- varImportance %>% mutate(Rank = paste0('#', dense_rank(desc(Importance))))


ggplot(rankImportance, aes(x = reorder(Variables, Importance), 
    y = Importance, fill = Importance)) +
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
    hjust=0, vjust=0.55, size = 4, colour = 'red') +
  labs(x = 'Variables') +
  coord_flip() + 
  theme_few()

```

### The Prediction
```{r}
prediction <- predict(rf_model, test)

test$type <- prediction

ggplot(test, aes(x = hair_length, y = has_soul, shape = type, color = type)) + geom_point()

plot(rf_model, log="y")
varImpPlot(rf_model)
MDSplot(rf_model, train$type)
getTree(rf_model, 3, labelVar = TRUE)
```

```{r}
# Save the solution to a dataframe with two columns: PassengerId and Survived (prediction)
solution <- data.frame(id = test$id, type = test$type)

# Write the solution to file
write.csv(solution, file = 'rf_mod_Solution2.csv', row.names = F)
```


# Thrid attempt - pure svm


```{r}
# factor(type) ~ bone_length + rotting_flesh + hair_length + has_soul + color
svmModel <- svm(type ~ bone_length + rotting_flesh + hair_length + has_soul + color, data=train)

prediction <- predict(svmModel, test)

test$type <- prediction

ggplot(test, aes(x = hair_length, y = has_soul, shape = type, color = type)) + geom_point()

# Does not consider the color
svmModel <- svm(type ~ hair_length + has_soul, data=train)

prediction <- predict(svmModel, test)

test$type <- prediction

ggplot(test, aes(x = hair_length, y = has_soul, shape = type, color = type)) + geom_point()

```

```{r}
# Save the solution to a dataframe with two columns: PassengerId and Survived (prediction)
solution <- data.frame(id = test$id, type = test$type)

# Write the solution to file
write.csv(solution, file = 'rf_mod_Solution4.csv', row.names = F)
```
